---
title: "Q1"
author: "Brian Williams"
date: "19 March 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## (1)  Put your exam code in the author space above
The data analysed in this folder is a new backpain dataset.  Descriptions of the variables are listed in the file Variables_BackPain2.xlsx in the ExamResources folder. There is another file, called BACKPAIN_EXTRACT_GenericIndividualQ.pdf, which you may also find useful.
The data is in a Stata data file called BackPainData_1.dta, which is located in the ExamData folder.


Here are the variables summarised.

```{r readData, echo = FALSE, message = FALSE}
library(tidyverse)
library(haven)
df <- read_dta("../ExamData/BackPainData_1.dta")
summary(df)
```

Take a look at the summary. There are a few un-helpful variable names, so make the following changes.

(2) Rename the variables:

1. quintile_c to wealthq
2. heightvalid to height
3. screen12
4. q4008
5. q4009 (See questionaire BACKPAIN_EXTRACT....pdf in ExamResources)

We will focus on the following 11 categorical variables. .:

1. residence
2. sex
3. country
4. wealthq (changed from quintile_c) 
5. agegr
6. educ4gp
7. bmi_4gp
8. drinkYN
9. arthritis
10. depression
11. eworked

When the df is loaded these 11 variables are listed as numeric.  

(2) Change the 11 listed variables to factors.

(3) The coding of the level names in these factors is not self-explanatory, so referring to the Variables_BackPain2.xlsx file, give concise, useful names to the levels of the 11 factor variables in the list.

(4) Add a summary with the new names and levels to this document.


(5) Prepare some boxplots and histograms

(5) Prepare some tables highlighting associations between any groupings of the variables above that you think will be of interest. You should prepare at least 6 groups and no more than 12. A three-way table is a single group and a two-way table is a single group. Do not use four-way grouping, please! Comment **briefly - one sentence** on each table.

(6) Prepare a scatter-plot of WHOQoL vs BPburden, with a linear smoother. Use the alpha transparency parameter in the plot. Repeat with WHOQoL vs your re-named screen12.
Comment on the comparison.

(7)


Let's select some variables of particular interest:

```{r selectVariables}
library(dplyr)
LM1 <- lm(screen12~BPburden, data = df)

```

```{r}
p <- ggplot(df, aes(x=BPburden, y = WHOQoL))+
  geom_point(alpha = 0.1)+
  stat_smooth(method = lm, lwd = 2)
p
```


```{r}
df <- df %>% mutate(sex = as.factor(sex)) %>% 
              mutate(depression = as.factor(depression)) %>% 
              mutate(eworked = as.factor(eworked)) %>% 
              mutate(country = as.factor(country)) %>% 
              filter(complete.cases(sex, depression, eworked, country, bmi))
p <- ggplot(df)+
  geom_boxplot(aes(x = eworked, y = bmi, fill = sex)) + ylim(0,70) +
  facet_grid(country~depression)
p
```

