---
title: "Q1"
author: "Brian Williams"
date: "19 March 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

(1)Put your exam code in the author space above

The data analysed in this folder is a new backpain dataset.  Descriptions of the variables are listed in the file Variables_BackPain2.xlsx in the ExamResources folder. There is another file, called BACKPAIN_EXTRACT_GenericIndividualQ.pdf, which you may also find useful.
The data is in a Stata data file called BackPainData_1.dta, which is located in the ExamData folder.


Here are the variables summarised.

```{r readData, echo = FALSE, message = FALSE}
library(tidyverse)
library(haven)
df <- read_dta("../Data/BackPainData_1.dta")
#summary(df)
```

Take a look at the summary. There are a few un-helpful variable names, so make the following changes.

(2)Rename the variables:

1. quintile_c to wealthq
2. heightvalid to height
3. screen12
4. q4008
5. q4009 (See questionaire BACKPAIN_EXTRACT....pdf in ExamResources)

```{r reNAme}
df <- df %>% rename(wealthq = quintile_c,
                    height = heightvalid,
                    disabIndex = screen12,
                    bP30 = q4008)
select(df, wealthq, height, disabIndex, bP30)


```


We will focus on the following 11 categorical variables. .:

1. residence
2. sex
3. country
4. wealthq (changed from quintile_c) 
5. agegr
6. educ4gp
7. bmi_4gp
8. drinkYN
9. arthritis
10. depression
11. eworked

When the df is loaded these 11 variables are listed as numeric.  

(3)Change the 11 listed variables to factors.

```{r asFactor}
df <- df %>% mutate_at(vars(residence, sex, country,
                       wealthq, agegr, educ4gp,
                       bmi_4gp, drinkYN, arthritis,
                       depression, eworked),
                       funs(as.factor))
summary(select(df, residence, sex, country,
                       wealthq, agegr, educ4gp,
                       bmi_4gp, drinkYN, arthritis,
                       depression, eworked ))

```


(4)The coding of the level names in these factors is not self-explanatory, so referring to the Variables_BackPain2.xlsx file, give concise, useful names to the levels of the 11 factor variables in the list.

```{r recode}
levels(df$residence) <- c("urban", "rural")
levels(df$sex) <- c("male", "female")
levels(df$country) <- c("China", "Ghana" ,"India", "Mexico", "Russian Fed", "Sth Africa")
levels(df$wealthq) <- c("poorest", "poor", "middle", "rich", "richest")
levels(df$agegr) <- c("50-59","60-69","70-79","80+ years")
levels(df$educ4gp) <- c("no primary", "primary", "secondary/high", "college/university")
levels(df$bmi_4gp) <- c("normal", "underweight","pre-obese", "obese")
levels(df$drinkYN) <- c("abstainers", "drinkers")
levels(df$arthritis) <- c("no", "yes")
levels(df$depression) <- c("no", "yes")
levels(df$eworked) <- c("currently working", "currently not working", "never worked")

```


(5)Add a summary with the new names and levels to this document.
```{r summary}
df %>% select(residence, sex, country,
                       wealthq, agegr, educ4gp,
                       bmi_4gp, drinkYN, arthritis,
                       depression, eworked) %>% 
  glimpse()
```


(6)Prepare a boxplot of bmi for men and women on the same plot and histograms of bmi for men and women on another plot. Put the two plots alongside one another.

```{r boxhist}
library(gridExtra)

p1 <- df %>% 
  ggplot()+
  aes(x = sex, y=bmi, fill=sex)+
  geom_boxplot()+
  theme_classic()

p2 <- df %>% 
  ggplot()+
  aes(x=bmi, fill=sex)+
  geom_histogram(binwidth = 0.5, position = "identity", alpha=.5) + xlim(0,75)

grid.arrange(p1,p2, ncol=2)

```


(7)Prepare some tables highlighting associations between any groupings of the categorical variables above that you think will be of interest. You should prepare at least 6 groups and no more than 8. A three-way table is a single group and a two-way table is a single group. Do not use four-way grouping, please! Comment **briefly - one sentence** on each table.

Note that 3-Way tables don't produce headers for the individual tables!

```{r tables1}
library(janitor)
library(knitr)
df %>%
  tabyl(residence, wealthq, show_na = F) %>% 
   adorn_percentages("all") %>% 
   adorn_rounding(2) %>% 
  kable()


```

```{r tables2}
df %>% 
  tabyl(wealthq, drinkYN, country, show_na = F) %>% 
   adorn_totals(where = c("row", "col")) %>% 
   adorn_percentages("all") %>% 
   adorn_rounding(2) %>% 
  kable()

```

```{r tables3}

```


```{r tables4}

```


```{r tables5}

```


```{r tables6}

```


(8)Prepare a scatter-plot of WHOQoL vs BPburden, with a linear smoother. Use the alpha transparency parameter in the plot. Repeat with WHOQoL vs your re-named screen12.
Comment on the comparison.

```{r plotSmooth}
p1 <- df %>% 
  ggplot()+
  aes(x = BPburden, y = WHOQoL)+
  geom_point(alpha=.05)+
  geom_smooth(method = "lm")+
  coord_equal()

#Repeat with WHOQoL vs your re-named screen12.
p2 <- df %>% 
  ggplot()+
  aes(x = disabIndex, y = WHOQoL)+
  geom_point(alpha=.05)+
  geom_smooth(method = "lm")+
  coord_equal()

grid.arrange(p1,p2, ncol=2)
```


(9)Use the lm() function to do the regression of WHOQoL vs BPburden and then plot the diagnostics.
lm1 = lm(WHOQoL~BPburden)

```{r selectVariables, fig.width=8}
lm1 = lm(df$WHOQoL~df$BPburden)
plot(lm1)
```



```


(10)Show boxplots of the bmi of Urban, Chinese women for all wealth categories and use colour to show smokers and non-smokers.

```{r  boxBMICWUS}
df %>% filter(sex=="female", residence=="urban", country=="China")%>%
  mutate(tobaccoYN = as.factor(tobaccoYN)) %>% 
  mutate(tobaccoYN = fct_recode(tobaccoYN, "Non-smoking" = "0","Smoking" ="1" )) %>% 
filter(complete.cases(residence, sex, country, tobaccoYN, wealthq)) %>% 
  ggplot()+
  aes(x=wealthq, y = bmi, fill=tobaccoYN)+
  geom_boxplot()

```
(11)Save the cleaned up df in the ExamData folder so that it can be loaded in Q2.

```{r save}
#save(df, "../Data/BackPain.Rdata")
```


(12) Test that the file knits properly, then save it in the examRcode folder.

