---
title: "Breast Cancer analysis"
author: "Brian Williams"
date: "8 April 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This analysis uses one of the standard (old) datasets used for investigating the characteristics of breast tumours in relation to the diagnosis of malign or benign. The [web site](http://archive.ics.uci.edu/ml/machine-learning-databases/breast-cancer-wisconsin/) gives detailed information about the variables in the dataset.
 
The dataset is available directly from the package mlbench (a collection of datasets suitable for machine learning).  You will need to have installed the package to continue.

```{r loadBCdata}
library(tidyverse)
library(mlbench)
data("BreastCancer")
dfBC <- BreastCancer
rm(BreastCancer)
summary(dfBC)         # Notice types of variables (See Environment pane)
dfBC$Bare.nuclei      # NA's in this one
```

Now undertake some exploratory analysis, focussing on the variables Cl.thickness, Cell.size and Cell.shape in relation to Class (the diagnosis).
This exploratory analysis should include tabulation and plotting using ggplot2.
First some tables:
```{r Table1}

```


```{r  Table2}

```

and then some plots...

```{r ggplot1}

```



```{r ggplot2}

```

.. and then a logistic regression of the probability of malignancy vs Cl.thickness.  You will need to mutate the Cl.thickness variable in dfBC to numeric and also mutate dfBC to create a new numeric variable Malignant which uses values 0 and 1 in place of "benign" and "malignant" 
It will look in part like this:

```{r ClThNum}
dfBC <- dfBC %>% mutate(Cl.thickness = as.numeric(Cl.thickness))
```


 mutate(Malignant = ifelse(Class == "benign", 0,1)

Once you have these two mutations you can use the function *glm*  with *family = binomial* to do the regression. 

```{r mutateMalig}
dfBC = dfBC %>% mutate(Malignant = ifelse(Class == "benign", 0,1))
```


Finally, you can pipe the mutated dfBC into ggplot: 

```{r NeedsMutatedDF}
dfBC %>%  ggplot(aes(x=Cl.thickness, y = Malignant)) +
  geom_jitter(height = 0.04, width = 0.2, alpha = 0.3, colour = "brown") +
  geom_smooth(method = "glm", 
              method.args = list(family = "binomial"))

```

You might like to tidy up the axis labels and give the plot a title?
                                    
                                