---
title: "Q1"
author: "Brian Williams"
date: "19 March 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

(1)Put your exam code in the author space above

The data analysed in this folder is a new backpain dataset.  Descriptions of the variables are listed in the file Variables_BackPain2.xlsx in the ExamResources folder. There is another file, called BACKPAIN_EXTRACT_GenericIndividualQ.pdf, which you may also find useful.
The data is in a Stata data file called BackPainData_1.dta, which is located in the ExamData folder.


Here are the variables summarised.

```{r readData, echo = FALSE, message = FALSE}
library(tidyverse)
library(haven)
df <- read_dta("../ExamData/BackPainData_1.dta")
#summary(df)
```

Take a look at the summary. There are a few un-helpful variable names, so make the following changes.

(2)Rename the variables:

1. quintile_c to wealthq
2. heightvalid to height
3. screen12
4. q4008
5. q4009 (See questionaire BACKPAIN_EXTRACT....pdf in ExamResources)

We will focus on the following 11 categorical variables. .:

1. residence
2. sex
3. country
4. wealthq (changed from quintile_c) 
5. agegr
6. educ4gp
7. bmi_4gp
8. drinkYN
9. arthritis
10. depression
11. eworked

When the df is loaded these 11 variables are listed as numeric.  

(3)Change the 11 listed variables to factors.

```{r rename}

```


(4)The coding of the level names in these factors is not self-explanatory, so referring to the Variables_BackPain2.xlsx file, give concise, useful names to the levels of the 11 factor variables in the list.

```{r recode}

```


(5)Add a summary with the new names and levels to this document.
```{r summary}

```


(6)Prepare a boxplot of bmi for men and women on the same plot and histograms of bmi for men and women on another plot. Put the two plots alongside one another.

```{r boxhist}

```


(7)Prepare some tables highlighting associations between any groupings of the categorical variables above that you think will be of interest. You should prepare at least 6 groups and no more than 8. A three-way table is a single group and a two-way table is a single group. Do not use four-way grouping, please! Comment **briefly - one sentence** on each table.

```{r tables1}

```

```{r tables2}

```

```{r tables3}

```


```{r tables4}

```


```{r tables5}

```


```{r tables6}

```


(8)Prepare a scatter-plot of WHOQoL vs BPburden, with a linear smoother. Use the alpha transparency parameter in the plot. Repeat with WHOQoL vs your re-named screen12.
Comment on the comparison.

```{r plotSmooth}

```


(9)Use the lm() function to do the regression of WHOQoL vs BPburden and then plot the diagnostics.


```{r selectVariables}
library(dplyr)


```


(10)Show boxplots of the bmi of Urban, Chinese women for all wealth categories and use colour to show smokers and non-smokers.

```{r  boxBMICWWS}

```
(11)Save the cleaned up df in the ExamData folder so that it can be loaded in Q2.

```{r save}

```


(12) Test that the file knits properly, then save it in the examRcode folder.

